# Accesing the data

## Before downloading datasets

If you are going to download data, you have to read the [Code of Conduct](https://docs.tradestatistics.io/index.html#code-of-conduct) first.

## API

*Since Dec 2021 the API features an interactive page https://api.tradestatistics.io/__docs__/.*

The advantage of the API over https download is that you can filter what to obtain and also access some additional tables.

To obtain exactly the same data as with compressed files, please refer to \@ref(yrpc-year-reporter-partner-and-commodity-code).

If you use R you'll need `jsonlite` package.

```{r message=FALSE, warning=FALSE}
library(jsonlite)
```

These packages are also useful:

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
```

### Available tables

```{r, warning=FALSE, message=FALSE, cache = TRUE}
as_tibble(fromJSON("https://api.tradestatistics.io/tables"))
```

### Metadata

```{r, warning=FALSE, message=FALSE, cache = TRUE}
## Countries (no filter)
rda_countries <- "countries.rda"

if (!file.exists(rda_countries)) {
  countries <- as_tibble(fromJSON(
    "https://api.tradestatistics.io/countries"
  ))

  save(countries, file = rda_countries, compress = "xz")

  countries
} else {
  load(rda_countries)

  countries
}

## Commodities (no filter)
rda_commodities <- "commodities.rda"

if (!file.exists(rda_commodities)) {
  commodities <- as_tibble(fromJSON(
    "https://api.tradestatistics.io/commodities"
  ))

  save(commodities, file = rda_commodities, compress = "xz")

  commodities
} else {
  load(rda_commodities)

  commodities
}
```

Please notice that these tables include some aliases. 

`countries` includes some meta-codes, `c-xx` where `xx` must the first two letters of a continent and `all`, this is:

```{r, echo = FALSE}
countries %>%
  filter(str_sub(country_iso, 1, 2) == "c-" | country_iso == "all") %>%
  select(country_iso, country_name_english) %>%
  rename(
    `Alias` = country_iso,
    `Meaning` = country_name_english
  )
```

`commodities` also includes some meta-codes, `xx` for the first two digits of a code and those digits are the commodity group and `all`, this is:

```{r, echo = FALSE}
commodities %>%
  filter(str_length(commodity_code) %in% 2:3) %>%
  select(commodity_code, commodity_fullname_english) %>%
  rename(
    `Alias` = commodity_code,
    `Meaning` = commodity_fullname_english
  )
```

### API parameters

The tables provided withing our API contain at least one of these fields:

* Year (`y`) 
* Reporter ISO (`r`)
* Partner ISO (`p`)
* Commodity Code (`c`)

The most detailed table is `yrpc` that contains all bilateral flows at commodity level.

With respect to `y` you can pass any integer contained in [1962,2020].

Both `r` and `p` accept any valid ISO code or alias contained in the [countries](https://api.tradestatistics.io/countries) table. For example, both `chl` (valid ISO code) and `c-am` (continent Americas, an alias) are valid API filtering parameters.

`c` takes any valid commodity code or alias from the [commodities](https://api.tradestatistics.io/commodities). For example, both `0101` (valid HS commodity code) and `01` (valid HS group code) are valid API filtering parameters.

By default the API takes `c = "all"` by default.

You can always skip `c`, but `y`, `r` and `p` are required to return data.

### Available reporters

The only applicable filter is by year.

```{r, warning=FALSE, message=FALSE, cache = TRUE}
# Available reporters (filter by year)
as_tibble(fromJSON(
  "https://api.tradestatistics.io/reporters?y=2018"
))
```

### YRPC (Year, Reporter, Partner and Commodity Code)

The applicable filters here are year, reporter, partner and commodity code.

```{r yrpc, warning=FALSE, message=FALSE, cache = TRUE}
# Year - Reporter - Partner - Commodity Code

yrpc <- as_tibble(fromJSON(
  "https://api.tradestatistics.io/yrpc?y=2018&r=can&p=usa&c=010121"
))

yrpc
```

Columns definition:

* `reporter_iso`: Official ISO-3 code for the reporter (e.g. the country that reports X dollars in exports/imports from/to country Y)
* `partner_iso`: Official ISO-3 code for the partner
* `commodity_code`: Official Harmonized System rev. 1992 (HS92) commodity code (e.g. according to the \code{commodities} table in the API, 8703 stands for "Motor cars and other motor vehicles; principally designed for the transport of persons (other than those of heading no. 8702), including station wagons and racing cars")
* `export_value_usd`: Exports measured in nominal United States Dollars (USD)
* `import_value_usd`: Imports measured in nominal United States Dollars (USD)

### YRC (Year, Reporter and Commodity Code)

The only applicable filter is by year, reporter and commodity code.

```{r yrc, warning=FALSE, message=FALSE, cache = TRUE}
# Year - Reporter - Commodity Code

yrc <- as_tibble(fromJSON(
  "https://api.tradestatistics.io/yrc?y=2018&r=chl"
))

yrc
```

### YRP (Year, Reporter and Partner)

The only applicable filter is by year, reporter and partner.

```{r yrp, warning=FALSE, message=FALSE, cache = TRUE}
# Year - Reporter - Partner
yrp <- as_tibble(fromJSON(
  "https://api.tradestatistics.io/yrp?y=2018&r=chl&p=arg"
))

yrp
```

### YC (Year and Commodity Code)

The only applicable filter is by year and commodity code.

```{r yc, warning=FALSE, message=FALSE, cache = TRUE}
# Year - Commodity Code
yc <- as_tibble(fromJSON(
  "https://api.tradestatistics.io/yc?y=2018&c=0101"
))

yc
```

### YR (Year and Reporter)

The only applicable filter is by year and reporter.

```{r yr, warning=FALSE, message=FALSE, cache = TRUE}
## Year - Reporter
yr <- as_tibble(fromJSON(
  "https://api.tradestatistics.io/yr?y=2018&r=chl"
))

yr
```

### Imputed tables

Just append "-imputed" to the API calls. Example: `https://api.tradestatistics.io/yr-imputed?y=2018&r=chl`.

## R Package

To ease API using, we provide an [R Package](https://ropensci.github.io/tradestatistics/). This package is a part of [ROpenSci](https://ropensci.org/) and its documentation is available on a separate [pkgdown site](https://ropensci.github.io/tradestatistics/).

Here's what the package does:

```{r, echo = F, fig.cap="R package flow"}
knitr::include_graphics("fig/data-diagram.svg")
```

## Dashboard (beta)

To ease API using, we provide a [Shiny Dashboard](https://shiny.tradestatistics.io/) that is still under improvements.

## Apache Arrow datasets

All the arrow datasets used to run the API and dashboard can be downloaded from [DigitalOcean Spaces](https://tradestatistics.ams3.digitaloceanspaces.com). These can be downloaded, for example, with [rclone](https://rclone.org/) by running the command
```
rclone sync spaces:tradestatistics/hs92-visualization hs92-visualization
```

Check the md5sums included in the file "md5sums.txt" within the download.
<!-- update with: find ~/hs-rev1992-visualization/* -type f -print0 | xargs -0 md5sum > ~/md5sums.txt -->
